  Q---
title: "pizza_codegrade"
author: "Joseph Christian"
date: "2022-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## library
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lm.beta))
install.packages("lm.beta")
```
##load data
```{r}
pizza <- suppressMessages(read_csv('pizza.csv'))
```


## instructions and beginning notes
```{r}
getwd()
describe(pizza)
pizza

```
## Question 1: Create a correlation matrix for temperature, bill, pizzas, and got_wine.  
```{r}
Q1 <- pizza %>%
  select(temperature, bill, pizzas, got_wine) %>%
  cor() %>%
  round(digits = 2)
```
  
## Q2: Create a correlation matrix of the relationships between time, temperature, bill, and pizzas for Laura in the East branch. 
```{r}
Q2 <- pizza %>%
  filter(operator == 'Laura' & branch == 'East') %>%
  select(time, temperature, bill, pizzas) %>%
  cor() %>%
  round(digits = 2)
```

## Q3: Run a regression predicting whether or not wine was ordered from temperature, bill, and pizza. 
```{r}
wineMod <- lm(got_wine ~ temperature + bill + pizzas, data = pizza)
summary(wineMod)
Q3 <- summary(wineMod)$coefficients %>%
  round(digits = 2)

#potentially log reg

```

## Q4: Run a regression predicting bill from temperature, pizzas, and got_wine.
```{r}
billMod <- lm(bill ~ temperature + pizzas + got_wine, data = pizza)
coe <- lm.beta(billMod)
Q4 <- coe$standardized.coefficients[2:4]

```

## Q5: Add operator to the regression from Q4.  Which is the better model?  
```{r}
pizza2 <- pizza %>%
  mutate(operator_bin = ifelse(operator == 'Laura', 1,0))
billMod2 <- update(billMod, .~. + operator_bin, data = pizza2)
Q5 <- AIC(billMod, k=2)
Q5
```






