---
title: "Untitled"
author: "Joseph Christian"
date: "2023-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("TTR")
```
```{r}
library(tidyverse)
```
```{r}
setwd("C:/Users/jbcwi/OneDrive/Desktop/DataScienceMS/DTSC560_DataScienceForBusiness")
```

##-----##
##Part1##
##-----##
```{r}
afdf <- read.csv("airfare.csv")
View(afdf)
```

```{r}
#create a time series plot showing yearly net revenue in billions
ggplot(afdf, aes(x=Year, y=Airfare)) +
  geom_line() +
  geom_point() +
  labs(title="Airfare vs Year", x="Year", y="Airfare")
```
##Quiz question #1: What patterns do you notice in the time series plot of average annual airfare? (MC)


##Quiz question #2: What is the 2019 forecast for average airfare using the simple moving average technique?
#353.760

##Quiz question #3: What is the MAE for the simple moving average method?
#19.301

##Quiz question #4: What is the 2019 forecast for average airfare using simple exponential smoothing with a smoothing constant of 0.2?
#376.636

##Quiz question #5: What is the MAE for the simple exponential smoothing method with asmoothing constant of 0.2?
#12.149

##Quiz question #6: What is the 2019 forecast for average airfare using simple exponential smoothing with a smoothing constant of 0.8?
#346.490

##Quiz question #7: What is the MAE for the simple exponential smoothing method with a smoothing constant of 0.8?
#14.362

##Quiz question #8: Reviewing the accuracy measure values for each of these forecasting attempts, which would you trust most to provide the most accurate forecast of 2019 average airfare?
#simple exponential smoothing with a smoothing constant of 0.2



##-----##
##Part2##
##-----##

```{r}
wmdf <- read.csv("warner_music.csv")
View(wmdf)
```

```{r}
#create a time series plot showing yearly net revenue in billions
ggplot(wmdf, aes(x=Quarter, y=Revenue)) +
  geom_line(group=1) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Revenue vs Quarter", x="Qrtr", y="Revenue")
```

##Quiz question #9: What patterns do you notice in the time series plot? (MC)
#upward horizontal trend and seasonality

```{r}
#add a column of consecutive numbers corresponding with each year
wmdf$Time <- 1:nrow(wmdf)
```

```{r}
#use simple linear regression analysis to create a regression equation for forecasting
wm_reg <- lm(Revenue ~ Time, data=wmdf)
summary(wm_reg)
```

##Quiz question #10: After conducting a linear regression to model the trend ONLY, what is the slope coefficient for the trend variable?
#23.190


```{r}
#create functions for the accuracy measures
mae <- function(actual, pred){
  mae <- mean(abs(actual-pred), na.rm=TRUE)
  return (mae)
}

mse <- function(actual, pred){
  mse <- mean((actual-pred)^2, na.rm=TRUE)
  return (mse)
}

rmse <- function(actual, pred){
  rmse <- sqrt(mean((actual-pred)^2, na.rm=TRUE))
  return (rmse)
}

mape <- function(actual, pred){
  mape <- mean(abs((actual-pred)/actual), na.rm=TRUE) * 100
  return (mape)
}
```

```{r}
wm_pred <- predict(wm_reg)
#calculate accuracy measures
mae(wmdf$Revenue, wm_pred)
mse(wmdf$Revenue, wm_pred)
rmse(wmdf$Revenue, wm_pred)
mape(wmdf$Revenue, wm_pred)
```

##Quiz question #11: What is the RMSE based on the results of the linear regression you just conducted?
#72.072

```{r}
#create dummy variables corresponding to each quarter
wmdf$Q1 <- ifelse(grepl("Q1", wmdf$Quarter),1,0)
wmdf$Q2 <- ifelse(grepl("Q2", wmdf$Quarter),1,0)
wmdf$Q3 <- ifelse(grepl("Q3", wmdf$Quarter),1,0)
wmdf$Q4 <- ifelse(grepl("Q4", wmdf$Quarter),1,0)
```

```{r}
#use multiple regression with quarter variables to generate a regression equation for forecasting
wm_reg2 <- lm(Revenue ~ Time + Q2 + Q3 + Q4, data=wmdf)
summary(wm_reg2)
```
##Quiz question #12: After conducting a linear regression to model both trend and seasonality, how should the regression coefficient for Quarter 2 be interpreted? (MC)
#You would subtract 110.83 from Q1


```{r}
wm_pred2 <- predict(wm_reg2)
#calculate accuracy measures
mae(wmdf$Revenue, wm_pred2)
mse(wmdf$Revenue, wm_pred2)
rmse(wmdf$Revenue, wm_pred2)
mape(wmdf$Revenue, wm_pred2)
```

##Quiz question #13: What is the RMSE based on the results of the second linear regression you just conducted?
#48.407

##Quiz question #14: What conclusion can you draw from comparing the accuracy measure values from the first regression to those from the second regression? (MC)
#they worked better combined


```{r}
#Predict NYT Revenue for 2022, Q1, Q2, Q3, Q4
#create an object with the time periods to use for the prediction
new <- data.frame(Time = c(26,27,28,29), Q2 = c(0,1,0,0), Q3 = c(0,0,1,0), Q4 = c(0,0,0,1))
predict(wm_reg2, newdata = new)
```

##Quiz question #15: Based on the second regression analysis modeling both trend and seasonality, what is the forecasted revenue for Warner Music Group for Quarter 4 of 2022?
#       1        2        3        4 
#1434.072 1323.906 1348.906 1405.559 

```{r}
#predict starbucks revenue for 2022
2.547 + 1.209*20
#predict starbucks revenue for 2023
2.547 + 1.209*21
#predict starbucks revenue for 2024
2.547 + 1.209*22
```

```{r}
#create an object with the time periods to use for the prediction
new <- data.frame(Time = c(20,21,22))
predict(sbreg, newdata = new)
```




```{r}
#create a vector of predicted values generated from the regression above
sb_pred = predict(sbreg)
```

```{r}
#calculate accuracy measures
mae(starbucksdf$NetRevenue, sb_pred)
mse(starbucksdf$NetRevenue, sb_pred)
rmse(starbucksdf$NetRevenue, sb_pred)
mape(starbucksdf$NetRevenue, sb_pred)
``` 

#steps to create a scatterplot of residuals vs predicted values of the dependent variable
```{r}
#look at residuals from time series regression
#steps to create a scatterplot of residuals vs predicted values of the dependent variable
#create a vectorof residuals generated from the regression above
sb_res = resid(sbreg)

#create a df of the predicted values and the residuals
pred_res_df <- data.frame(sb_pred, sb_res)

#create a scatterplot of the residuals versus the predicted values
ggplot(pred_res_df, aes(x=sb_pred, y=sb_res)) +
  geom_point() +
  labs(title = "Plot of residuals vs. Predicted values", x = "predicted values", y = "Residuals")
```


##Seasonality
##Multiple and single regression
##New York Times data
```{r}
nytdf <- read.csv("NYT_revenue.csv")
View(nytdf)
```

```{r}
#create a time series plot showing yearly net revenue in billions
ggplot(nytdf, aes(x=Quarter, y=Revenue)) +
  geom_line(group=1) +
  geom_point() +
  labs(title="NYT Qrtly Revenue (2013-2016)", x="Quarter", y="Revenue")
```

```{r}
#create dummy variables corresponding to each quarter
nytdf$Q1 <- ifelse(grepl("Q1", nytdf$Quarter),1,0)
nytdf$Q2 <- ifelse(grepl("Q2", nytdf$Quarter),1,0)
nytdf$Q3 <- ifelse(grepl("Q3", nytdf$Quarter),1,0)
nytdf$Q4 <- ifelse(grepl("Q4", nytdf$Quarter),1,0)
```

```{r}
#use multiple regression with quarter variables to generate a regression equation for forecasting
nytreg <- lm(Revenue ~ Q1 + Q2 + Q3, data=nytdf)
summary(nytreg)
```

```{r}
#Predict NYT Revenue for 2017, Q1, Q2, Q3, Q4
#create an object with the time periods to use for the prediction
new <- data.frame(Q1 = c(1,0,0,0), Q2 = c(0,1,0,0), Q3 = c(0,0,1,0))
predict(nytreg, newdata = new)
```



##Horizontal trend, seasonality
##Multiple and Linear Regression
##Whole Foods Data
```{r}
wfdf <- read.csv("Whole_foods.csv")
```

```{r}
#create a time series plot showing quarterly net sales
ggplot(wfdf, aes(x=Quarter, y=Net.Sales)) +
  geom_line(group=1) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Whole Foods Qrtrly Net Sales (2005-2016)(in millions)", x = "Quarter", y = "Net Sales")
```

```{r}
#add a column of consecutive numbers corresponding with each year
wfdf$Time <- 1:nrow(wfdf)

#create dummy variables corresponding to each quarter
wfdf$Q1 <- ifelse(grepl("Q1", wfdf$Quarter),1,0)
wfdf$Q2 <- ifelse(grepl("Q2", wfdf$Quarter),1,0)
wfdf$Q3 <- ifelse(grepl("Q3", wfdf$Quarter),1,0)
wfdf$Q4 <- ifelse(grepl("Q4", wfdf$Quarter),1,0)
```

```{r}
#use regression with the time variable to generate a regression equation for forecasting
wfreg <- lm(Net.Sales ~ Time, data = wfdf)

wf_pred <- predict(wfreg)
```

```{r}
#calculate accuracy measures
mae(wfdf$Net.Sales, wf_pred)
mse(wfdf$Net.Sales, wf_pred)
rmse(wfdf$Net.Sales, wf_pred)
mape(wfdf$Net.Sales, wf_pred)
```

```{r}
#use multiple regression with quarter variables to generate a regression equation for forecasting
wfreg2 <- lm(Net.Sales ~ Time + Q2 + Q3 + Q4, data=wfdf)
summary(wfreg2)
```

```{r}
 #create a vector of predicted values genereated from the multiple regression above
wf_pred2 <- predict(wfreg2)
```

```{r}
#calculate accuracy measures
mae(wfdf$Net.Sales, wf_pred2)
mse(wfdf$Net.Sales, wf_pred2)
rmse(wfdf$Net.Sales, wf_pred2)
mape(wfdf$Net.Sales, wf_pred2)
```

```{r}

#create an object with the time periods to use for the prediction
new <- data.frame(Time = c(49,50,51,52), Q2 = c(0,1,0,0), Q3 = c(0,0,1,0), Q4 = c(0,0,0,1))
predict(wfreg2, newdata = new)
```


##Quadratic Models
##LinkedIn data
```{r}
lidf <- read.csv("linked_in.csv")
```

```{r}
#create a time series plot showing number of linked in members by qrtr in millions
ggplot(lidf, aes(x = Quarter, y = Members)) +
  geom_line(group=1) +
  geom_point() +
  theme(axis.text.x = element_text( angle = 90)) +
  labs(title = "LinkedIn members by Qrtr (millions) (2009-2014)", x = "Quarter", y = "Members")
```

```{r}
#add a column of consecutive numbers corresponding with each quarter
lidf$Time <- 1:nrow(lidf)
```

```{r}
#use simple linear regression analysis to create a regression equation for forecasting
lireg <- lm(Members ~ Time, data=lidf)
summary(lidf)
```

```{r}
#create a vector of predicted valyues generated from the regression above
li_pred = predict(lireg)

#calculate accuracy measures
mae(lidf$Members, li_pred)
mse(lidf$Members, li_pred)
rmse(lidf$Members, li_pred)
mape(lidf$Members, li_pred)
```

```{r}
#create a new variable that squares the Time variable
lidf$Time2 <- lidf$Time^2
```

```{r}
#use a quadratic regression model to create a regression equation for forecasting
liregquad <- lm(Members ~ Time + Time2, data = lidf)
summary(liregquad)
```

```{r}
#create a vector of predicted valyues generated from the regression above
li_pred2 = predict(liregquad)

#calculate accuracy measures
mae(lidf$Members, li_pred2)
mse(lidf$Members, li_pred2)
rmse(lidf$Members, li_pred2)
mape(lidf$Members, li_pred2)
```

```{r}
#predict linkedin membership for qrtr 3 and 4 in 2014
#create an object with the tie periods to use for the prediction
new <- data.frame(Time = c(23, 24), Time2 = c(529, 576))
predict(liregquad, newdata = new)
```

